---
title: "Core Faculty Model Evaluation - Data Exploration"
format: html
execute:
  echo: false
---

```{r, include = FALSE}
library(dplyr)
library(stringr)
library(tidyr)
library(readxl)
library(htmltools)
library(ggplot2)
require(kableExtra)

data <- read_excel(
  "data/BIDMC_Med_Neuro_SPE_Comments_Dataset_07242025.xlsx",
  1
) |>
  mutate(
    ROTATION_DATE = as.Date(ROTATION_DATE, format = "%m/%d/%Y"),
    qID = as.factor(QUESTION) |> as.integer(),
    qSummative = qID < 3
  )
dict <- read_excel("data/BIDMC_Med_Neuro_SPE_Comments_Dataset_07242025.xlsx", 2)
```

# About the data
This dataset was collected from the Harvard Medical School's medical education
program clinical rotations and contains the written student 
evaluations by the Faculty.

<blockquote>NOTE: This document does not contain any sensitive information, only summary 
statistics</blockquote>

# Basic info about the data

```{r}
tags$ul(
  tags$li("Number of students:", n_distinct(data$LEARNER_ANON_ID)),
  tags$li("Number of faculty:", n_distinct(data$EVALUATOR_ID)),
  tags$li(
    "Clerkships evaluated:",
    paste(unique(data$CLERKSHIP) |> sort(), collapse = ", ")
  ),
  tags$li(
    "Hospital locations:",
    paste(unique(data$LOCATION) |> sort(), collapse = ", ")
  ),
  tags$li(
    "Academic Years:",
    paste(unique(data$ACAD_YR) |> sort(), collapse = ", ")
  ),
  tags$li(
    sprintf(
      "Total number of evaluations: %i of which %i are summaries",
      n_distinct(data$ROWID),
      sum(data$SUMMARY_FLG == "Y")
    )
  )
)
```

The data is stored in an Excel file that contain the following column information:
```{r, results = 'asis'}
half <- ceiling(length(dict$DEFINITION) / 2)

tags$table(tags$tr(
  tags$td(tags$ul(lapply(dict$DEFINITION[1:half], tags$li))),
  tags$td(tags$ul(lapply(
    dict$DEFINITION[(half + 1):length(dict$DEFINITION)],
    tags$li
  )))
))
```

# Student demographics
```{r}
data |>
  group_by(LEARNER_ANON_ID) |>
  slice(1) |>
  group_by(GENDER, ACAD_PROG_TRK) |>
  summarise(
    n = n(),
    mean_age = sprintf(
      "%.0f (\u00B1%.1f)",
      mean(AGE),
      ifelse(is.na(sd(AGE)), 0, sd(AGE))
    ),
    .groups = "drop"
  ) |>
  mutate(
    n = sprintf(
      "%i (%.1f%%)",
      n,
      n / sum(n) * 100
    )
  ) |>
  kableExtra::kable()
```

# By year
```{r}
data |>
  group_by(ACAD_YR) |>
  summarise(
    dates = sprintf("%s - %s", min(ROTATION_DATE), max(ROTATION_DATE)),
    nStudents = n_distinct(LEARNER_ANON_ID),
    nEvals = n(),
    nSummaries = sum(SUMMARY_FLG == "Y"),
    .groups = "drop"
  ) |>
  kableExtra::kable()
```

# Data distribution
## Number of evaluations per student for completed academic years
```{r}
result <- data |>
  group_by(LEARNER_ANON_ID, CLERKSHIP, ACAD_PROG_TRK) |>
  summarise(
    ACAD_YR = unique(ACAD_YR),
    nEval = n(),
    nSummary = n_distinct(SUMMARY_FLG),
    nFaculty = n_distinct(EVALUATOR_ID),
    .groups = "drop"
  )

ggplot(
  result |> filter(ACAD_YR < max(ACAD_YR)),
  aes(x = nEval, fill = CLERKSHIP)
) +
  geom_histogram(binwidth = 1, color = "white") +
  labs(
    title = NULL,
    x = "Number of evaluations"
  ) +
  theme_minimal()

```

## Average number of evaluations per faculty for a completed academic year
```{r}
result <- data |>
  filter(ACAD_YR < max(ACAD_YR)) |>
  group_by(ACAD_YR, EVALUATOR_ID) |>
  summarise(
    nStudents = n_distinct(LEARNER_ANON_ID),
    nEvals = n() / n_distinct(ACAD_YR),
    nSummary = n_distinct(SUMMARY_FLG),
    .groups = "drop"
  )

ggplot(result, aes(x = nEvals)) +
  geom_histogram(bins = 50, color = "white") +
  scale_x_log10(breaks = c(1, 2, 3, 5, 10, 25, 50, 100, max(result$nEvals))) +
  labs(
    title = NULL,
    x = "Number of evaluations  (log scale!)"
  ) +
  theme_minimal()
```

# Written Feedback Statistics

The following table list the questions used as a prompt for the evaluations and a 
few basic metrics.

<blockquote>NOTE: The first two question are only answered by faculty who do a summative evaluation</blockquote>

```{r}
data |>
  mutate(
    nChar = nchar(ANSWER_TXT),
    nWords = str_count(ANSWER_TXT, "[\\w+\\-']+")
  ) |>
  group_by(QUESTION) |>
  summarise(
    n = n(),
    `average_nChar` = sprintf(
      "%.0f (\u00B1%.0f)",
      mean(nChar),
      sd(nChar)
    ),
    `average_nWords` = sprintf(
      "%.0f (\u00B1%.0f)",
      mean(nWords),
      sd(nWords)
    ),
    .groups = "drop"
  ) |>
  arrange(n) |>
  kableExtra::kable()
```

## Number of questions completed per faculty member for a student
```{r}
result <- data |>
  filter(ACAD_YR < max(ACAD_YR)) |>
  group_by(qSummative, LEARNER_ANON_ID, EVALUATOR_ID) |>
  summarise(nAnswered = n_distinct(qID), .groups = "drop")

group_labels <- c(
  "TRUE" = "Summative eval",
  "FALSE" = "Regular eval"
)

ggplot(result, aes(x = nAnswered)) +
  geom_histogram(binwidth = 1, color = "white") +
  facet_wrap(
    ~qSummative,
    scales = "free",
    labeller = as_labeller(group_labels)
  ) +
  scale_x_continuous(breaks = function(x) floor(min(x)):ceiling(max(x))) +
  labs(
    title = NULL,
    x = "Number of questions"
  ) +
  theme_minimal()
```

# Most common words
Below is a list of the 30 most commonly used words longer than 5 characters
```{r}
words <- data$ANSWER_TXT |>
  str_extract_all("[\\w+\\-']+") |>
  unlist() |>
  tolower() |>
  table()

words <- data.frame(
  words = names(words),
  n = as.integer(words)
)

multiCol <- function(df, n_col, n_row) {
  if (exists("n_row")) {
    df <- df[1:(min(n_row * n_col, nrow(df))), ]
  } else {
    n_row <- ceiling(nrow(df) / n_col)
  }

  df <- lapply(seq(1, nrow(df), by = n_row), function(i) {
    df[i:(i + n_row - 1), ]
  })

  do.call(cbind, df)
}

words <- words |> filter(nchar(words) > 5) |> arrange(desc(n))

multiCol(words, n_col = 3, n_row = 10) |> kable()
```

